---
title: "Moneyball in the NBA"
output: html_notebook
---

## Load data and examine structure

```{r}
NBA = read.csv("NBA_train.csv")
str(NBA)
```

Creating a table comparing season wins to playoffs variable. From this, it looks like getting around 42 wins gets you a pretty good chance of making it to the playoffs.

```{r}
table(NBA$W, NBA$Playoffs)

```

Creating variable "PTSdiff" and creating a scatter plot with Wins

```{r}
NBA$PTSdiff = NBA$PTS - NBA$oppPTS
plot(NBA$PTSdiff, NBA$W)
```

Creating linear model to predict Wins using the variable PTSdiff

```{r}
WinsReg = lm(W~PTSdiff, data = NBA)
summary(WinsReg)
```

Our regression equation is: *Wins = 41 + 0.0326(PTSdiff)*

What does PTSdiff need to equal in order to get at least 42 wins, and therefore make it to the playoffs?

*PTSdiff \>= (42-41)/0.0326 = 30.67*

So we need to score on average 31 more points in order to win at least 42 games.

Creating a linear regression to predict points using 2 pointer attempts, 3 pointer attempts, free throw attempts, assists, offensive rebounds, defensive rebounds, turnovers, steals, and blocks.

```{r}
PointsReg = lm(PTS ~ X2PA + X3PA + FTA + AST + ORB + DRB + TOV + STL + BLK, data = NBA)
summary(PointsReg)
```

```{r}
SSE = sum(PointsReg$residuals^2)
RMSE = sqrt(SSE/nrow(NBA))
RMSE
```

This is the Root Mean Square Error, which tells us we make an error of about 184.4 season points.

Removing insignificant variables, staring with TOV because it has the highest P value.

```{r}
PointsReg2 = lm(PTS ~ X2PA + X3PA + FTA + AST + ORB + DRB + STL + BLK, data = NBA)
summary(PointsReg2)
```

```{r}
PointsReg3 = lm(PTS ~ X2PA + X3PA + FTA + AST + ORB + STL + BLK, data = NBA)
summary(PointsReg3)
```

```{r}
PointsReg4 = lm(PTS ~ X2PA + X3PA + FTA + AST + ORB + STL, data = NBA)
summary(PointsReg4)
```

```{r}
SSE_4 = sum(PointsReg4$residuals^2)
RMSE_4 = sqrt(SSE_4/nrow(NBA))
RMSE_4
```

So although r\^2 increased, RMSE is about the same. This is still the best model.

Loading in the test file.

```{r}
NBA_test = read.csv("NBA_test.csv")
```

Calculating Out-of-sample R2

```{r}
PointsPrediction = predict(PointsReg4, newdata = NBA_test)
SSE = sum((PointsPrediction - NBA_test$PTS)^2)
SST = sum((mean(NBA$PTS) - NBA_test$PTS)^2)
R2 = 1 - SSE/SST
R2
```

Calculating RMSE

```{r}
RMSE = sqrt(SSE/nrow(NBA_test))
RMSE
```
