---
title: "R Notebook"
output: html_notebook
---
Loading data. Need to include stringAsFactors in order for R to interpret chr format as factors
```{r}
train = read.csv("pisa2009train.csv", stringsAsFactors = TRUE)
test = read.csv("pisa2009test.csv", stringsAsFactors = TRUE)
```

Finding avg reading test score of males
```{r}
tapply(train$readingScore, train$male, mean)
```

Removing null values
```{r}
train = na.omit(train)
test = na.omit(test)
```

Setting reference level of the factor variable "raceeth" to "White" because it is the most common value for that variable in the dataset. R by default sets the reference level to the first value alphabetically, in this case "American Indian / Alaska Native"
```{r}
train$raceeth = relevel(train$raceeth, "White")
test$raceeth = relevel(test$raceeth, "White")
```

Building model. Note "." means all
```{r}
lmScore = lm(readingScore ~ ., data = train)
summary(lmScore)
```
RMSE
```{r}
SSE = sum(lmScore$residuals^2)
RMSE = sqrt(SSE/nrow(train))
RMSE
```

Score prediction
```{r}
predTest = predict(lmScore, newdata = test)
summary(predTest)
```

RMSE
```{r}
SSE = sum((predTest - test$readingScore)^2)
RMSE = sqrt(SSE/nrow(test))
RMSE
```


R2 of predictions
```{r}
SST = sum((mean(train$readingScore) - test$readingScore)^2)
R2 = 1 - SSE/SST
R2
```

